%------------------------
% Curriculum Vitae (CV)
% Aravinda Raman Jatavallabha
% Comprehensive | PhD | Research | O-1A / EB-1A
% All content from https://aravinda-1402.github.io/ — no fabrication
%------------------------

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage[usenames,dvipsnames]{color}
\usepackage{enumitem}
\usepackage[pdftex,colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.4in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-0.45in}
\addtolength{\textheight}{1in}
\urlstyle{rm}
\raggedright
\setlength{\parindent}{0pt}

\titleformat{\section}{\vspace{-8pt}\scshape\raggedright\large\bfseries}{}{0em}{}[\color{black}\titlerule \vspace{-4pt}]

\newcommand{\cvEntry}[4]{\textbf{#1} \hfill #2 \\ \textit{#3} \hfill \textit{#4} \vspace{2pt}}
\newcommand{\cvItem}[1]{\item \small #1 \vspace{-1pt}}

%-----------------------------
\begin{document}
%-----------------------------

%---------- HEADER ----------
\begin{center}
{\LARGE \textbf{Aravinda Raman Jatavallabha}} \\
Raleigh, NC, United States \\
AI Software Engineer \& Independent Researcher \\[4pt]
\href{mailto:aravindaraman14@gmail.com}{aravindaraman14@gmail.com} \,$|$\, \href{tel:+19193270958}{+1 (919) 327-0958} \,$|$\, \href{https://aravinda-1402.github.io/}{Portfolio} \,$|$\, \href{https://scholar.google.com/citations?user=fXRLPrEAAAAJ\&hl=en}{Google Scholar} \,$|$\, \href{https://linkedin.com/in/aravinda-jatavallabha}{LinkedIn} \,$|$\, \href{https://github.com/aravinda-1402}{GitHub}
\end{center}
\vspace{4pt}

%---------- RESEARCH INTERESTS ----------
\section{Research Interests}
\small
LLMs and RAG; privacy-aware ML systems; dynamic/streaming NLP; temporal graph learning; time-series forecasting; trustworthy AI (fairness, compliance).
\vspace{4pt}

%---------- RESEARCH SUMMARY ----------
\section{Research Summary}
\small
My research focuses on the design and deployment of robust, privacy-aware, and scalable machine learning systems, with emphasis on large language models (LLMs), retrieval-augmented generation (RAG), and dynamic representation learning for evolving data streams. I have contributed to both theoretical and applied research across NLP, graph learning, and time-series modeling, with publications in IEEE, Springer, and arXiv venues.

My work spans healthcare AI (document understanding, clinical summarization, compliance-aware LLM pipelines), dynamic language modeling (incremental BERT, temporal GNNs), and applied ML systems in public safety, networking, and recommender systems. Several of these contributions have been deployed in production environments or evaluated on large-scale real-world datasets.
\vspace{4pt}

%---------- EDUCATION ----------
\section{Education}

\cvEntry{Master of Computer Science (Data Science (Specialization))}{Aug 2023--May 2025}{North Carolina State University, Raleigh, NC}{GPA: 4.0/4.0 \href{https://drive.google.com/file/d/1uPDeOb_QWkxuiRlXxEfEIW2JagjP5olQ/view?usp=drive_link}{[Credential]}}
\\
\textbf{Relevant coursework:} Neural Networks and Deep Learning; Foundations of Data Science; Automated Learning and Data Analysis; Database Management Systems; Machine Learning with Graphs; Natural Language Processing; Generative AI; Privacy in Artificial Intelligence; Design and Analysis of Algorithms.
\vspace{6pt}

\cvEntry{Bachelor of Technology in Information Technology (Big Data Analytics (Minor))}{Jun 2019--Jul 2023}{Manipal Institute of Technology, Manipal, India}{GPA: 8.64/10.0 \href{https://drive.google.com/file/d/1egeuz65fwpKCLLV0Bi9-n2kIEI_nJkYz/view?usp=sharing}{[Credential]}}
\\
\textbf{Relevant coursework:} Data Mining; Machine Learning; Pattern Recognition; Algorithms; Big Data Analytics; Data Structures.
\vspace{6pt}

%---------- PUBLICATIONS \& PREPRINTS ----------
\section{Publications \& Preprints}

\subsection*{Peer-Reviewed Publications}
\begin{enumerate}[leftmargin=*,nosep]
  \item \textbf{TrustBench: Benchmarking Trustworthy Large Language Models with Cost and Stability Metrics.} IEEE SoutheastCon 2026 (Accepted). Reproducible benchmark for deployment trust in LLMs: desired behavior rate, run stability, confidence dispersion, and cost-effectiveness across GPT-4.1, GPT-4o, GPT-5, Claude Sonnet 4.5, and Grok 4.1; evaluates ambiguity handling, safety, and multi-step repeatability.
  \item \textbf{Learning Dynamic Representations in Large Language Models for Evolving Data Streams.} ICPR 2024, Springer. \href{https://link.springer.com/chapter/10.1007/978-3-031-78169-8_16}{[Link]} Dynamic representation learning in LLMs using graph-based attention; incremental BERT (iBERT) and dynamic graphs for streaming text; iDCWE. (Contributing author; primary responsibility for model design and experiments.)
  \item \textbf{SDN-Based Multipath Data Offloading Scheme Using Link Quality Prediction for LTE and WiFi Networks.} IEEE. \href{https://ieeexplore.ieee.org/abstract/document/10767139}{[Link]} LSTM/BLSTM for link quality (RSSI, PDR); 2.1\% better prediction; 6.29\% HetNet throughput improvement. (Contributing author; primary responsibility for model design and experiments.)
  \item \textbf{Diabetes Prognosis using Machine Learning: A Comparative Analysis of Classification Algorithms.} IEEE. \href{https://ieeexplore.ieee.org/abstract/document/9985487}{[Link]} KNN, Random Forest, ANN; 87.89\% (Random Forest); preprocessing, feature engineering.
\end{enumerate}
\vspace{4pt}

\subsection*{Preprints and Manuscripts Under Review}
\begin{enumerate}[leftmargin=*,nosep]
  \item \textbf{Engineering Deployment-Aware Fairness Auditing for Healthcare AI in Cloud Environments.} Under Review. Comprehensive technical analysis of fairness auditing in healthcare ML: subgroup evaluation, counterfactual testing, lifecycle-based auditing, feedback loops, distributional shift, regulatory compliance, and cloud-native MLOps pipelines for continuous fairness monitoring.
  \item \textbf{Using Transformer-Based Models to Optimize Inventory Replenishment Decisions in Dynamic E-Commerce Markets.} Under Review. Multi-head attention for demand forecasting in e-commerce; 23.7\% reduction in stockouts, 18.4\% decrease in excess inventory costs; captures seasonality, promotional lift, and product substitution patterns.
  \item \textbf{Privacy Awareness in Large Language Models: Input Regurgitation and Prompt-Induced Sanitization for HIPAA and GDPR Compliance.} Under Review. Input regurgitation risk in LLMs; prompt-induced sanitization strategies for privacy-sensitive deployments under HIPAA and GDPR.
  \item \textbf{Prompting for LLM Security and RAG: A Survey from Zero-Shot to Automatic Prompt Optimization (APO) and Prompt-Injection Defenses.} Under Review. Survey of prompting strategies and security defenses for prompt injection and RAG systems; zero-shot to APO.
  \item \textbf{Dynamic Graph Representation Learning using Temporal and Topological Information.} Under Review. TDGNN: time-aware message passing, graph topology, point process theory for social and interaction networks.
  \item \textbf{Improving Fairness in Visual Recognition through Feature Distillation and Adversarial Debiasing.} \href{https://drive.google.com/file/d/1R5COOj2JY-Omxbo9tsHxiVuBQ6Jpqou2/view?usp=drive_link}{[Link]} MMD-based distillation from unfair teacher; feature invariance across protected attributes; reduced disparity, competitive accuracy.
  \item \textbf{Multimodal Conversation Derailment Detection: An Integrated Framework for Early Risk Assessment.} \href{https://drive.google.com/file/d/18lczqZVlpkBnT68dv-w-8IUNeHbGWiB4/view?usp=drive_link}{[Link]} Hierarchical transformer; BERT + Faster R-CNN; 71.0\% accuracy, 78.3\% AUC; +6.1\% over text-only.
  \item \textbf{Graph Contrastive Learning for Optimizing Sparse Data in Recommender Systems with LightGCL.} \href{https://drive.google.com/file/d/1ND3R-cwxMfOw16Wutt6KToq9feTAH5JE/view?usp=drive_link}{[Link]} SVD-based augmentation; structural refinement and global collaborative relation modeling; strong on sparsity and item-popularity bias.
  \item \textbf{Tesla's Autopilot: Ethical Implications and Policy Considerations in Autonomous Vehicle Systems.} arXiv:2409.17380. \href{https://arxiv.org/abs/2409.17380}{[Link]} Seven-step ethical decision-making; user behavior, system constraints, regulatory implications; framework for safety-critical deployments.
  \item \textbf{Deciphering Air Travel Disruptions: A Machine Learning Approach to Flight Delay Prediction.} arXiv:2408.02802. \href{https://arxiv.org/abs/2408.02802}{[Link]} LSTM, Hybrid LSTM, Bi-LSTM vs regression; departure time, airline, weather.
  \item \textbf{Pediatric Bone Age Assessment using Deep Learning Models: A Comparative Study of CNN Architectures.} arXiv:2207.10169. \href{https://arxiv.org/abs/2207.10169}{[Link]} VGG-16, InceptionV3, XceptionNet, MobileNet on pediatric X-rays; skeletal maturity; MAE-based comparison.
\end{enumerate}
\vspace{2pt}

%---------- RESEARCH EXPERIENCE ----------
\section{Research Experience}

\cvEntry{Machine Learning Researcher}{Jan 2023--Jun 2023}{Defence Research and Development Organisation (DRDO)}{Bengaluru, Karnataka, India}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{Proposed and implemented a Temporal Graph Neural Network (TGNN) at the Centre for AI \& Robotics (CAIR) Lab using PyTorch Geometric (PyG) to model dynamic user interactions over time for semantic-aware behavioral forecasting; improved prediction accuracy by 2\% compared to then-existing baselines.}
  \cvItem{Developed a streaming NLP pipeline leveraging BERT for learning contextualized word embeddings incrementally, addressing semantic drift by dynamically updating token representations without full retraining.}
  \cvItem{Pioneered the iBERT + Temporal GNN hybrid, integrating graph temporal reasoning with evolving language representations; real-time semantic understanding; reduced model retraining latency by 40\%. Validated on Reddit, arXiv, news, Yelp, and Ciao.}
  \cvItem{Achieved language model perplexity of 3.19, surpassing the prior benchmark of 3.40 on masked language prediction across evolving news article streams.}
\end{itemize}
\vspace{4pt}

\cvEntry{Machine Learning Researcher}{Mar 2021--Jun 2022}{Manipal Institute of Technology}{Udupi, Karnataka, India}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{Bone Age Assessment: medical imaging of pediatric patients (1--288 months); VGG-16, MobileNet, InceptionV3, XceptionNet; MAE-based comparative analysis.}
  \cvItem{LSTM and Bi-LSTM for channel link quality prediction in IoT network traces; 99.4\% accuracy.}
  \cvItem{Automated ETL, model training, REST API deployment via Apache Airflow, AWS SageMaker, ECR, Flask; 60\% reduction in manual intervention.}
  \cvItem{AWS S3 for data ingestion and versioning; Keras, TensorFlow, Matplotlib for modeling and visualization.}
  \cvItem{Collaborated with faculty on technical implementations at the intersection of healthcare, networking, and AI.}
\end{itemize}
\vspace{4pt}

%---------- INDUSTRY EXPERIENCE ----------
\section{Industry Experience}

\cvEntry{Software Engineer --- AI Platforms \& Full-Stack Systems}{Jun 2025--Present}{Long Health}{San Jose, California, United States · Remote}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{Designed and deployed AWS Lambda + ECS + S3 serverless workflows and internal REST/event-driven APIs for document ingestion, metadata indexing, and retrieval; 10K+ healthcare documents weekly; 35\% throughput gain.}
  \cvItem{Implemented RabbitMQ-based OCR (AWS Textract for scanned paperwork), RAG, LLM summarization, and ICD-10 inference pipelines backed by ChromaDB; 99.9\% uptime; monitoring and alerting for pipeline health and latency.}
  \cvItem{Integrated OpenAI Whisper for real-time transcription of physician--patient conversations and auto-generation of structured physical exam and summary reports.}
  \cvItem{Integrated OpenAI LLMs for real-time summaries, structured extraction, medical Q\&A, and physician chatbot; optimized RAG chunking and retrieval to curb hallucinations; 50\% reduction in documentation and triage time.}
  \cvItem{Partnered with clinical and compliance teams to ship HIPAA- and PHI/PII-safe workflows. Built and maintained Angular/NestJS physician portals with role-aware dashboards.}
\end{itemize}
\vspace{4pt}

\cvEntry{Data Science Intern / Co-op --- ML \& Predictive Analytics}{May 2024--Jun 2025}{SmartProtect Public Safety Solutions}{Wilmington, Delaware, United States · Remote}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{Improved 911 call-volume forecasting accuracy by 20\% via A/B/n-tested ARIMA, Prophet, LSTM, and Logistic Regression on AWS SageMaker.}
  \cvItem{Built scalable ETL with Apache Airflow, Flask APIs, and Snowflake; 1.2M+ dispatch records across regions.}
  \cvItem{Deployed drift detection and auto-retraining with AWS Lambda and S3 versioning; 35\% faster ML updates.}
  \cvItem{Engineered predictive staffing algorithms from historical and forecasted volume; 18\% overtime reduction, 22\% workforce utilization gain.}
  \cvItem{Developed full-stack internal dashboard (Spring Boot + Angular) with RBAC for shift recommendations, anomalies, and forecast trends. Integrated Azure OpenAI and LLaMA via LangChain for real-time dispatcher transcript analysis and anomaly detection.}
\end{itemize}
\vspace{4pt}

\cvEntry{Data Science Intern --- ML \& Analytics}{May 2022--Jul 2022}{Merkle}{Bengaluru, Karnataka, India · Hybrid}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{Led a team of 4: XGBoost, LightGBM, LSTM on 10M+ retail records; 10\% uplift in campaign profitability for Home Depot via revenue optimization.}
  \cvItem{Engineered ETL with PySpark, SQL, Snowflake; 16M+ rows; partitioning, caching, indexing; 40\% query latency improvement.}
  \cvItem{Integrated LLM-based vector embeddings to cluster product descriptions; personalized targeting and recommendation precision.}
  \cvItem{Automated batch pipelines (PySpark, SQL) for ingestion, transformation, validation; pricing analytics. Power BI and Tableau dashboards for pricing, segmentation, and sales KPIs.}
  \cvItem{Collaborated with product engineers and analysts to integrate model outputs into APIs and UIs; A/B testing.}
\end{itemize}
\vspace{4pt}

%---------- TEACHING \& MENTORSHIP ----------
\section{Teaching \& Mentorship}

\cvEntry{Graduate Teaching Assistant}{Aug 2024--May 2025}{North Carolina State University}{Raleigh, North Carolina, United States}
\begin{itemize}[leftmargin=*,nosep]
  \cvItem{\textbf{CSC 522: Automated Learning and Data Analysis} (Jan--May 2025, Prof. Thomas Price). Reviewed and evaluated student assignments and projects on data mining, supervised and unsupervised learning, predictive modeling, and Python-based ML workflows (scikit-learn). Provided detailed feedback on pipelines, train/validation/test splits, evaluation metrics, regression, classification, and clustering (K-Means, DBSCAN, hierarchical). Held office hours; addressed questions on feature engineering, cross-validation, and model selection.}
  \cvItem{\textbf{CSC 591/791: Real-Time AI and Machine Learning Systems} (Aug--Dec 2024, Prof. Xipeng Shen). Graded assignments, projects, and exams on real-time AI systems, optimization (TensorRT, ONNX Runtime), and performance trade-offs in ML deployment. Supported students on latency optimization, model compression, quantization, and efficient inference (batch vs. streaming). Assisted with debugging and best practices for production ML systems.}
  \cvItem{Collaborated with faculty on academic standards, course operations, and exam design. Managed course materials, rubrics, and student queries; developed and applied grading rubrics for consistent evaluation across assignments and projects.}
\end{itemize}
\vspace{4pt}

%---------- TECHNICAL PROJECTS ----------
\section{Technical Projects}

\begin{itemize}[leftmargin=*,nosep]
  \cvItem{\textbf{CoveredAI --- Health Insurance Document Assistant.} Full-stack AI app using LLMs and RAG to simplify health insurance documents. Users upload plans, ask natural-language questions, view smart summaries, compare plans side-by-side, and export personalized PDF reports. End-to-end semantic search, summarization, and secure document handling. Tech: React, TypeScript, TailwindCSS, Flask, LangChain, OpenAI GPT, FAISS. \href{https://github.com/aravinda-1402/coveredai-healthclaim-rag-agent}{[GitHub]}}
  \cvItem{\textbf{Legal Query AI Assistant.} Chatbot with LLMs (OpenAI GPT / LLaMA) and RAG to retrieve, summarize, and answer legal queries from document repositories. Vector-based search for high accuracy and fast retrieval. Tech: GPT, LangChain, RAG, Vector Search. \href{https://github.com/aravinda-1402/Legal-Query-AI-Assistant}{[GitHub]}}
  \cvItem{\textbf{Lane Detection for Autonomous Vehicles.} Hybrid SegNet + LSTM to detect lane lines, compute lane curvature, and measure vehicle offset from dashcam or video input. OpenCV-based image preprocessing and real-time inference. Tech: SegNet, LSTM, OpenCV. \href{https://github.com/aravinda-1402/Lane-Detection-for-Autonomous-Vehicles}{[GitHub]}}
  \cvItem{\textbf{Store Demand Forecasting using Time-Series and Neural Networks.} Hybrid CNN + BiLSTM to predict daily item-level sales across 10 stores. Kaggle Store-Item Demand (2013--2017). Benchmark against XGBoost, ANN, and ARIMA; lowest MSE. Supports retail inventory and replenishment decisions. Tech: CNN, BiLSTM, XGBoost, Time Series. \href{https://github.com/aravinda-1402/Store-Demand-Forecasting-using-Time-Series-and-Neural-Networks}{[GitHub]}}
  \cvItem{\textbf{Image-to-Image Translation using CycleGAN.} Unpaired domain translation without paired datasets: Monet paintings to real photographs, human faces to zombies. Trained on publicly available datasets; real-time inference. Unsupervised generative learning for computer vision. Tech: CycleGAN, PyTorch, Computer Vision. \href{https://github.com/aravinda-1402/CycleGAN-for-Image-Translation}{[GitHub]}}
  \cvItem{\textbf{Brain Tumor Segmentation using MRI Scans.} U-Net for semantic segmentation on brain MRI; pixel-level mask prediction for tumor regions. LGG MRI Segmentation (Kaggle), FLAIR sequences. High segmentation accuracy and visual interpretability for medical diagnostics. Tech: U-Net, TensorFlow/Keras, Medical Imaging. \href{https://github.com/aravinda-1402/Brain-MRI-Classification-Segmentation}{[GitHub]}}
  \cvItem{\textbf{Customer Churn Prediction.} Full ML pipeline: Apache Airflow for orchestration, AWS (S3, SageMaker, ECR) for training and storage, Dockerized Flask APIs for serving. Scalable deployment and real-time churn inference. Tech: AWS, Airflow, Docker, Flask, scikit-learn. \href{https://github.com/aravinda-1402/Customer-Churn-Prediction}{[GitHub]}}
  \cvItem{\textbf{Cold Email Generator.} Automated outreach: LangChain ChatGroq + LLaMA3 with ChromaDB to extract job descriptions, match user skills via embeddings, and generate personalized cold emails using RAG. Streamlit UI for job-to-email generation. Tech: LangChain, LLaMA3, ChromaDB, RAG, Streamlit. \href{https://github.com/aravinda-1402/Cold_Email_Generator}{[GitHub]}}
\end{itemize}
\vspace{2pt}

%---------- HONORS, AWARDS, \& MEMBERSHIPS ----------
\section{Honors, Awards, \& Memberships}

\begin{itemize}[leftmargin=*,nosep]
  \cvItem{\textbf{Sigma Xi, The Scientific Research Honor Society} --- Full Member (Jan 2026--Present). Elected through peer nomination and review based on demonstrated original research contributions. Interdisciplinary research honor society; \href{https://www.sigmaxi.org/}{sigmaxi.org}.}
  \cvItem{\textbf{Stellar Team Award}, Merkle --- Top-performing project in the internship cohort. Recognized for innovation, measurable impact (campaign profitability, ETL, ML modeling), and cross-functional execution.}
\end{itemize}
\vspace{2pt}

%---------- TECHNICAL SKILLS ----------
\section{Technical Skills}

\textbf{Programming \& scripting:} Python, SQL, HTML/CSS, JavaScript, TypeScript, Angular, NestJS, Flask, REST API, Git.

\textbf{Data handling \& visualization:} Pandas, NumPy, SciPy, OpenCV, Matplotlib, Seaborn, Plotly, Power BI, Tableau, Excel, Document AI \& OCR.

\textbf{Machine learning \& deep learning:} Supervised \& Unsupervised Learning; Scikit-learn, PyTorch, TensorFlow, Keras; NLTK, SpaCy, PyG. Models: CNNs, RNNs, LSTMs, GNNs, Recommender Systems, Anomaly Detection, Ranking Models, Time Series Forecasting.

\textbf{Generative AI \& LLMs:} LLMs, LangChain, RAG, Prompt Engineering, OpenAI API, Hugging Face, LLaMA, Pinecone, ChromaDB.

\textbf{Cloud, MLOps \& big data:} AWS (Lambda, ECS, S3, ECR, SageMaker), Azure, Apache Spark, Hadoop, Snowflake, Apache Airflow, Serverless Architectures, RabbitMQ, Docker.

\textbf{Statistics \& experimentation:} Statistical Modeling, Hypothesis Testing, A/B Testing, Regression, Classification, Clustering, Time Series Analysis, HIPAA \& PHI Compliance.
\vspace{4pt}

%---------- PROFESSIONAL CERTIFICATIONS ----------
\section{Professional Certifications}

\begin{itemize}[leftmargin=*,nosep]
  \cvItem{\textbf{Data Science Specialization}, NC State University. \href{https://drive.google.com/file/d/1-o0rSSCy6qN3lY99xNvIWSQEKPD0ELD7/view?usp=sharing}{[Certificate]}}
  \cvItem{\textbf{Machine Learning Specialization}, Stanford University (Coursera). \href{https://www.coursera.org/account/accomplishments/verify/9AB6K4JPZSQ9}{[Verify]}}
  \cvItem{\textbf{Deep Learning Specialization}, DeepLearning.AI (Coursera). \href{https://www.coursera.org/account/accomplishments/specialization/DFFQB96C6FJ2}{[Verify]}}
  \cvItem{\textbf{AI Summer School}, IIIT Hyderabad. \href{https://drive.google.com/file/d/1jPhjHO8WfoukW9IKUp8yKpRyoLKbnjFB/view}{[Certificate]}}
\end{itemize}
\vspace{4pt}

\end{document}
